#Lab_05  {lab_05}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and janitor.

**Task** load these two packages.

```{r}
# Turn off scientific notation
options(scipen=999)

library(tidyverse)
library(janitor)
```

### Load Life Expectancy Data

You'll need to load :

CDC_Life_Census_Tract_2010_2015.csv

It is included in your lab_05.zip file.

About the data: 
Life Expectancy at Birth for U.S. States and [Census Tracts, 2010-2015](https://www.cdc.gov/nchs/data-visualization/life-expectancy/index.html) 

It shows estimates of U.S. life expectancy at birth by state and census tract for the period 2010-2015. Estimates were produced for 67,148 census tracts, covering all states and the District of Columbia (D.C.). These estimates come from the National Center for Health Statistics (NCHS), the National Association for Public Health Statistics and Information Systems (NAPHSIS), and the Robert Wood Johnson Foundation (RWJF).


**Task** Create a codeblock below this task, then load the data sets and assign them to appropriate variable names. 

## Answer questions

**Q1.** In your lab_05 folder, there is a csv called CDC_Life_Census_Tract_2010_2015.csv.  It contains all of the life expectancy data for the United States.

You are going to examine this data and learn the range of death rates in Baltimore and then in Maryland in general.

Do the following:
-  Load tidyverse and janitor
-  Load the data CDC_Life_Census_Tract_2010_2015.csv into a dataframe called us_death. 
Include the code to show the number of records. Clean the names using Janitor
-  Filter this according to Maryland into second table called md_death
-  Filter this according to Baltimore City into a third table called balt_death

**A1.**

```{r}
us_death <- read_csv("CDC_Life_Census_Tract_2010_2015.csv") %>% 
  as.data.frame()
  clean_names %>% 
  nrow()
  
md_death <- us_death %>% 
  filter(State == "Maryland")

balt_death <- us_death %>% 
  filter(County == "Baltimore city, MD")

```

**Q2.** 
For this question, load a modified dataset that has the same information but with the Baltimore neighborhoods added. It's called balt_death_census.csv. Call this table balt_death2. 
Produce a summary of basic statistics on the average life expectancy rate.

Show the neighborhoods with a life expectancy rate below the 1st quartile, arranged in ascending order. 


**A2.** 
```{r}
balt_death2 <- read.csv("balt_death_census.csv") %>% 
  as.data.frame()

balt_death2 %>% 
  summary(life_expectancy)

balt_death2 %>% 
  clean_names %>% 
  filter(life_expectancy < 69.58) %>% 
  arrange(life_expectancy) 


```

**Q3.** Use md_death, the Maryland life expectancy rates.
Produce summary statistics of the state of Maryland life expectancy rate. 
Produce a table with the 20 communities with the lowest life expectancy. 
Produce a table with the 20 communities with the highest life expectancy. 

Then use us_death, and determine the average life expectancy nationwide.

Then write 150-200 words on this document below your code that compares your findings of Baltimore, the state and nation. 

**A3.**

```{r}

md_death %>% 
  summary(life_expectancy)

min_expectancy <- md_death %>% 
  clean_names %>% 
  slice_min(life_expectancy, n = 20)

max_expectancy <- md_death %>% 
  clean_names %>% 
  slice_max(life_expectancy, n = 20)

us_death %>% 
  summary(life_expectancy)

```
From this data I found out that the state of Maryland as a whole has a mean and median life expectancy that is close to the total U.S. life expectancy. However, it is clear that the life expectancy in Baltimore varies wildly from this because almost all of the minimum life expectancy rates are neighborhoods in Baltimore. Unfortunately from all I have learned about wealth inequality and the lack of resources and cleanliness in poor areas of Baltimore that have been historically segregated and made so that minorities neighborhoods struggle more, known as the "Black Butterfly", this makes sense and shows the discrepancy between Baltimore city and the rest of the state. 


**Q4.**

Use balt_death. Use the case_when command in mutate to categorize all Baltimore City Census tracts as being above or below the city's average life expectancy rate. 
Then produce a short table that summarizes the number of Census tracts above or below the average life expectancy rate. Write two sentences summarizing your findings.


```{r}
balt_death %>% 
  summary(life_expectancy)

balt_more_less <- balt_death %>% 
  clean_names %>% 
  mutate(
    more_less = case_when(
      life_expectancy == 'NA' ~ "NA",
      life_expectancy > '73.04' ~ "MORE",
      life_expectancy < '73.04' ~ "LESS"
    )
  )

balt_more_less %>% 
  clean_names %>% 
  select(county, census_tract_number, life_expectancy, more_less) 
```
Based on the table I created, I can see that mostly Baltimore neighborhoods in the 2700 range Census Tracts are above the average life expectancy, meaning they are likely the more wealthy neighborhoods. Those in the 1600 range Census Tracts seem to have the most number of areas that are below the average life expectancy, indicating they are likely poorer areas.

