#lab_08
Sean Mussenden, Aug 16, 2022

## About this lab

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Load janitor and the tidyverse
```{r}
library(tidyverse)
library(janitor)
```

## Load data

**Task**: Load four tables of U.S. Census data.  Each table has one row per Maryland county with 2020 population information.  Each table has these same two columns:

* geoid - a unique identifier for each county
* name - the county name

The third column in each table represents the total population in one of four demographic categories.

* maryland_population.csv - total_pop - total population 
* maryland_hispanic.csv - hispanic_pop - total Hispanic population
* maryland_white.csv - white_pop - total White alone population
* maryland_black.csv - black_pop - total Black alone population

Some of the information in the tables may need to be cleaned or validated in order to produce the correct answer.  

```{r}
total_pop <- read_csv("data/maryland_population.csv") %>% 
  clean_names
hispanic_pop <- read_csv("data/maryland_hispanic.csv") %>% 
  clean_names
white_pop <- read_csv("data/maryland_white.csv") %>% 
  clean_names
black_pop <- read_csv("data/maryland_black.csv") %>% 
  clean_names

```


## Questions

**Q1**  What Maryland county has the largest Hispanic population (by percentage of total population)? 
**A1** Montgomery County, Maryland. 19.5% of the total population in the county is Hispanic.
```{r}
hispanic_pop$name <- str_replace(hispanic_pop$name,"Montgomery County,Maryland", "Montgomery County, Maryland")

hispanic_total <- total_pop %>% 
  as.data.frame(hispanic_total) %>% 
  clean_names %>% 
  group_by(geoid, name) %>% 
  inner_join(hispanic_pop, c("geoid" = "geoid", "name" = "name")) %>% 
  select(total_pop, hispanic_pop) %>% 
  mutate(pct_hispanic = (hispanic_pop / total_pop)*100) %>% 
  arrange(desc(pct_hispanic))

```

**Q2**  What is the total population of people who identify as white alone in Maryland? 
**A2**  The total population of people who identify as white alone in Maryland is 3,275,048 people.
```{r}

white_pop$white_pop <- str_replace(white_pop$white_pop, "13510000", "13510")

 white_pop %>% 
   as.data.frame() %>% 
   clean_names() %>% 
   mutate(white_pop = as.numeric(white_pop)) %>% 
   summarise(
    white_pop = sum(white_pop, na.rm=TRUE))

```

**Q3** What is the average Black population in all Maryland counties? 
**A3** The average Black population in all Maryland counties is 53,864.91 people.
```{r}
black_pop %>% 
  as.data.frame() %>% 
  clean_names() %>% 
  mutate(black_pop = as.numeric(black_pop)) %>% 
  summarise(
    black_pop = mean(black_pop, na.rm=TRUE)
    )
  

```
